---
title: "RTST Choking, Aspiration, and Pneumonia Trend Report"
format: docx
execute: 
  echo: false
  warning: false
params:
  census_path: ""
  cho_path: ""
  current_asp_path: ""
  past_asp_path: ""
  year: ""
  months: ""
---

```{r, setup}

library(tidyverse)
library(reshape2)
library(gt)
library(readxl)

source("scripts/process_data.R")


#path1.1 <- "E:/CHOKING Event Tracking.xlsm"
#path1.2 <- "E:/2023 - ASPIRATION - PNEUMONIA Event Tracking.xlsm"
#path1.2.2 <- "E:/2022 - ASPIRATION - PNEUMONIA Event Tracking.xlsm"
#path2 <- "E:/2023 Agency Census Listing.xlsx"
#year = "2023"
#month = c("10", "11", "12")


path1.1 <-  params$cho_path
path1.2 <- params$current_asp_path
path1.2 <- str_sub(path1.2, 2, nchar(path1.2) - 2)
path1.2.2 <- params$past_asp_path
path1.2.2 <- str_sub(path1.2.2, 2, nchar(path1.2.2) - 2)
path2 <- params$census_path
year = params$year
month = params$months

```

```{r, report range}
#| output: asis

if (month[1] == "NA") {
  
  print(
    str_glue("# {year} yearly choking, aspiration, pneumonia events")
    )
  
} else {
  
  quart <- tibble(month = month)
  
  quart$quart <- case_match(
    quart$month,
    "01" ~ "1",
    "02" ~ "1",
    "03" ~ "1",
    "04" ~ "2",
    "05" ~ "2",
    "06" ~ "2",
    "07" ~ "3",
    "08" ~ "3",
    "09" ~ "3",
    "10" ~ "4",
    "11" ~ "4",
    "12" ~ "4",
  )
  
  print(
    str_glue("# {year} Q{quart$quart[1]} choking, aspiration, pneumonia events")
    )
}

```

```{r, data import}
#log data import, choking
choking_data <- import_event_data(
  path_str = path1.1,
  event_str = "choking_events"
)

#collapse subagencies into the parent agency name
choking_data <- collapse_names(choking_data)


#log data import, aspiration
aspiration_data <- import_event_data(
  path_str = path1.2,
  event_str = "aspiration_events"
)

#collapse subagencies into the parent agency name
aspiration_data <- collapse_names(aspiration_data)


#log data import, pneumonia
pneumonia_data <- import_event_data(
  path_str = path1.2,
  event_str = "pneumonia_events"
)

#collapse subagencies into the parent agency name
pneumonia_data <- collapse_names(pneumonia_data)


##census data import
census_titles <- read_xlsx(path = ".titles/event_titles.xlsx",
          sheet = "census",
          skip = 0,
          col_types = c("text", "text", "text", "text"))

census <- read_xlsx(path = path2,
                     sheet = 1,
                     skip = 2,
                     col_names = census_titles$clean,
                     col_types = census_titles$type)

census <- change_names(census)

```

```{r, data prep}
#select data and filter for quarterly months
#choking
choking_data <- choking_data[,1:60]

choking_data_report <- quarterly_report_months(choking_data)

#aspiration
aspiration_data <- aspiration_data[,1:43]

aspiration_data_report <- quarterly_report_months(aspiration_data)

#pneumonia
pneumonia_data <- pneumonia_data[,1:14]

pneumonia_data_report <- quarterly_report_months(pneumonia_data)



#census data prep
key <- census_key(census)


#match census data to reportable events
choking_data_report <- match_key(choking_data_report, key)


aspiration_data_report <- match_key(aspiration_data_report, key)


pneumonia_data_report <- match_key(pneumonia_data_report, key)

```

## Choking trend report

Total choking events

```{r, total choking table}

choking_data_report_months <- choking_data_report |>
  count(inc_mon_rep, name = "event")

choking_data_report_months2 <- choking_data_report |>
  select(inc_mon_rep, inc_mon_num)

choking_data_report_months <- unique(left_join(choking_data_report_months, choking_data_report_months2) |>
  arrange(inc_mon_num) |>
  select(-inc_mon_num))

choking_data_report_months |>
  rbind(c("Total", nrow(choking_data_report))) |>
  gt(rowname_col = "inc_mon_rep") |>
  cols_label(event = "Count")

```

Affected age groups

```{r, choking ages}

choking_data_report |>
  count(age_cat, name = "event") |>
  gt(rowname_col = "age_cat") |>
  cols_label(event = "Count") 
```

Current therapies

```{r, choking current therapies}

choking_data_report |>
  count(current_services, name = "event") |>
  gt(rowname_col = "current_services") |>
  cols_label(event = "Count")
```

Referrals made

```{r, referral for choking}

choking_data_report |>
  count(referral_for, name = "choking") |>
  gt(rowname_col = "referral_for") |>
  cols_label(choking = "Count")
```

Incidences by agency

```{r, choking event by agency}

agency_count <- choking_data_report |>
  count(agency, name = "event") 

agency_census <- unique(choking_data_report |> select(agency, cen_mean))

agency_count <- left_join(agency_count, agency_census) 

#process factors
a_list <- str_split(choking_data_report$contrib_fact, 
          pattern = ",",
          n = Inf, 
          simplify = TRUE)

a_list <- data.frame(a_list)

a_list$agency <- choking_data_report$agency
a_list2 <- drop_na(melt(a_list, id.vars = "agency")) |>
  select(!variable)

a_list2$value <- str_trim(a_list2$value)

a_list3 <- a_list2 |>
  count(agency, value) |>
  filter(value != "") |>
  mutate(
    factor_report = paste0(value, "(", n, ")")
  )

a_list4 <- a_list3 |>
  group_by(agency) |>
  arrange(desc(n)) |>
  mutate(cont_fact_report = str_flatten(factor_report, collapse = ", "))

agency_report <- a_list4 |>
  select(agency, cont_fact_report)

agency_count <- unique(left_join(agency_count, agency_report)) |>
  mutate(percent = round(event/cen_mean*100, 1))

agency_count |>
  arrange(desc(percent)) |>
  select(agency, cen_mean, event, percent, cont_fact_report) |>
  gt(rowname_col = "agency") |>
  cols_label(
    cen_mean = "Census",
    event = "Events",
    percent = "Percent",
    cont_fact_report = "Contributing factors"
    )
```

Overall contributing factors

```{r, overall factors choking}

a_list4 |>
  group_by(value) |>
  count(value, name = "count") |>
  mutate(percent = round(count/nrow(choking_data_report)*100, 1)) |>
  arrange(desc(percent)) |> 
  ungroup() |>
  gt(rowname_col = "value") |>
  cols_label(
    count = "Count",
    percent = "Percent",
    )
```

## Aspiration trend report

Total aspiration events

```{r, aspiration total}

aspiration_data_report_months <- aspiration_data_report |>
  count(inc_mon_rep, name = "event")

aspiration_data_report_months2 <- aspiration_data_report |>
  select(inc_mon_rep, inc_mon_num)

aspiration_data_report_months <- unique(left_join(aspiration_data_report_months, aspiration_data_report_months2) |>
  arrange(inc_mon_num) |>
  select(-inc_mon_num))

if (nrow(aspiration_data_report_months) == 0) {
  print("No events were reported")
} else {
  aspiration_data_report_months |>
  rbind(c("Total", nrow(aspiration_data_report))) |>
  gt(rowname_col = "inc_mon_rep") |>
  cols_label(event = "Count")
}


```

Affected age groups

```{r, aspiration age groups}


aspiration_data_report |>
  count(age_cat, name = "event") |>
  gt(rowname_col = "age_cat") |>
  cols_label(event = "Count") 

```

Current therapies

```{r, aspiration therapies}

aspiration_data_report |>
  count(current_services, name = "event") |>
  gt(rowname_col = "current_services") |>
  cols_label(event = "Count")
```

Referrals made

```{r, aspiration referrals}

if (nrow(aspiration_data_report_months) == 0) {
  print("No events were reported")
} else {
  aspiration_data_report |>
  count(referral_for, name = "event") |>
  gt(rowname_col = "referral_for") |>
  cols_label(event = "Count")
}


```

Incidences by agency

```{r, aspiration event by agency}

agency_count_asp <- aspiration_data_report |>
  count(agency, name = "event") 

agency_census_asp <- unique(aspiration_data_report |> select(agency, cen_mean))

agency_count_asp <- left_join(agency_count_asp, agency_census_asp) 

if (nrow(aspiration_data_report_months) == 0) {
  print("No events were reported")
} else {
  #process factors
a_list_asp <- str_split(aspiration_data_report$contrib_fact, 
          pattern = ",",
          n = Inf, 
          simplify = TRUE)

a_list_asp <- data.frame(a_list_asp)

a_list_asp$agency <- aspiration_data_report$agency
a_list2_asp <- drop_na(melt(a_list_asp, id.vars = "agency")) |>
  select(!variable)

a_list2_asp$value <- str_trim(a_list2_asp$value)

a_list3_asp <- a_list2_asp |>
  count(agency, value) |>
  filter(value != "") |>
  mutate(
    factor_report = paste0(value, "(", n, ")")
  )

a_list4_asp <- a_list3_asp |>
  group_by(agency) |>
  arrange(desc(n)) |>
  mutate(cont_fact_report = str_flatten(factor_report, collapse = ", "))

agency_report_asp <- a_list4_asp |>
  select(agency, cont_fact_report)

agency_count_asp <- unique(left_join(agency_count_asp, agency_report_asp)) |>
  mutate(percent = round(event/cen_mean*100, 1))

agency_count_asp |>
  arrange(desc(percent)) |>
  select(agency, cen_mean, event, percent, cont_fact_report) |>
  gt(rowname_col = "agency") |>
  cols_label(
    cen_mean = "Census",
    event = "Events",
    percent = "Percent",
    cont_fact_report = "Contributing factors"
    )
}


```

Overall contributing factors

```{r, overall factors aspiration}

if (nrow(aspiration_data_report_months) == 0) {
  print("No events were reported")
} else {
  a_list4_asp |>
  group_by(value) |>
  count(value, name = "count") |>
  mutate(percent = round(count/nrow(choking_data_report)*100, 1)) |>
  arrange(desc(percent)) |> 
    ungroup() |>
  gt(rowname_col = "value") |>
  cols_label(
    count = "Count",
    percent = "Percent",
    )
}


```

## Pneumonia trend report

Total pneumonia events

```{r, pneumonia total}

pneumonia_data_report_months <- pneumonia_data_report |>
  count(inc_mon_rep, name = "event")

pneumonia_data_report_months2 <- pneumonia_data_report |>
  select(inc_mon_rep, inc_mon_num)

pneumonia_data_report_months <- unique(left_join(pneumonia_data_report_months, pneumonia_data_report_months2) |>
  arrange(inc_mon_num) |>
  select(-inc_mon_num))

if (nrow(pneumonia_data_report_months) == 0) {
  print("No events were reported")
} else {
  pneumonia_data_report_months |>
  rbind(c("Total", nrow(pneumonia_data_report))) |>
  gt(rowname_col = "inc_mon_rep") |>
  cols_label(event = "Count")
}
```

Affected age groups

```{r, pneumonia age groups}

pneumonia_data_report$age_cat <- cut_interval(pneumonia_data_report$age, length = 10, right = FALSE)

pneumonia_data_report$age_cat <- recode(pneumonia_data_report$age_cat,
                                   "[0,10)" = "20s",
                                   "[10,20)" = "20s",
                                   "[20,30)" = "20s",
                                   "[30,40)" = "30s",
                                   "[40,50)" = "40s",
                                   "[50,60)" = "50s",
                                   "[60,70)" = "60s",
                                   "[70,80)" = "70s",
                                   "[70,80]" = "70s",
                                   "[80,90)" = "80s",
                                   "[80,90]" = "80s",
                                   "[90,100)" = "80s",
                                   "[90,100]" = "80s")

pneumonia_data_report |>
  count(age_cat, name = "event") |>
  gt(rowname_col = "age_cat") |>
  cols_label(event = "Count") 
```

Repeat event dates of pneumonia and aspiration from names of people affected by pneumonia or aspiration this quarter.

The search range includes the current year to date (up to the current quarter) and the previous year (Q1-Q4 from last year).

```{r, pneumonia repeat episodes}

#names of people from this quarter
pneumonia_names <- unique(pneumonia_data_report |>
  select(first_name, last_name, dob))

aspiration_names <- unique(aspiration_data_report |>
  select(first_name, last_name, dob))

#remove duplicate names
pneumonia_names <- anti_join(pneumonia_names, aspiration_names)

combined_names <- bind_rows(pneumonia_names, aspiration_names)

combined_names_name <- as.data.frame(str_split(combined_names$last_name, pattern = ",", simplify = TRUE))

combined_names$last_name <- combined_names_name$V1

#events from year-to-date
inc_date <- as_tibble(str_split(pneumonia_data$event_date, pattern = "-", n = 3, simplify = T))
  
  pneumonia_data$inc_mon <- inc_date$V2

if (month[1] == "NA") {
  
  pneumonia_to_date <- pneumonia_data |>
    select(first_name, last_name, event_date)
  
} else {
  
  pneumonia_to_date <- pneumonia_data |> 
    mutate(inc_mon_num = as.numeric(inc_mon),
           event_date_char = as.character(event_date)) |>
    filter(inc_mon_num <= as.numeric(month[3])) |>
    select(first_name, last_name, event_date)

}

#import last year data
pneumonia_data_last_year <- import_event_data(
  path_str = path1.2.2,
  event_str = "pneumonia_events"
)


pneumonia_data_last_year$event_date <- as.Date(substr(pneumonia_data_last_year$event_date, 1, 10))

pneumonia_data_last_year <- pneumonia_data_last_year |>
  mutate(
    event_date_char = as.character(event_date)
  )

pneumonia_data_last_year_na <- is.na(pneumonia_data_last_year$last_name)

pneumonia_data_last_year <- pneumonia_data_last_year[!pneumonia_data_last_year_na,]

pneumonia_data_last_year <- pneumonia_data_last_year |>
  select(first_name, last_name, event_date)


##we also want to bring in the aspiration-pneumonia events
#events from year-to-date
inc_date <- as_tibble(str_split(aspiration_data$event_date, pattern = "-", n = 3, simplify = T))
  
  aspiration_data$inc_mon <- inc_date$V2

if (month[1] == "NA") {
  
  aspiration_to_date <- aspiration_data |> 
    select(first_name, last_name, event_date)
} else {
  
  aspiration_to_date <- aspiration_data |> 
    mutate(inc_mon_num = as.numeric(inc_mon),
           event_date_char = as.character(event_date)) |>
    filter(inc_mon_num <= as.numeric(month[3])) |>
    select(first_name, last_name, event_date)

}
  


if (nrow(aspiration_to_date) == 0) {
    print("There were no repeated events this year to date for people affected by aspiration or pneumonia this quarter")
}

#events from last year
aspiration_data_last_year <- import_event_data(
  path_str = path1.2.2,
  event_str = "aspiration_events"
)


aspiration_data_last_year$event_date <- as.Date(substr(aspiration_data_last_year$event_date, 1, 10))

aspiration_data_last_year_na <- is.na(aspiration_data_last_year$last_name)

aspiration_data_last_year <- aspiration_data_last_year[!aspiration_data_last_year_na,]

aspiration_data_last_year <- aspiration_data_last_year |>
  select(first_name, last_name, event_date)

if (nrow(aspiration_data_last_year) == 0) {
    print("There were no previous events last year for people affected by aspiration or pneumonia this quarter")
}


if (nrow(aspiration_to_date) > 0 & nrow(aspiration_data_last_year) > 0) {

  asp_pne_last_year_to_date <- bind_rows(aspiration_to_date, aspiration_data_last_year) |>
  mutate(event_date_char = paste0(event_date, " (asp)"))

  pneumonia_last_year_to_date <- bind_rows(pneumonia_to_date, pneumonia_data_last_year, asp_pne_last_year_to_date) |>
  mutate(event_date_char = as.character(event_date))

} else if (nrow(aspiration_to_date) > 0 & nrow(aspiration_data_last_year) == 0) {

  asp_pne_last_year_to_date <- bind_rows(aspiration_to_date) |>
  mutate(event_date_char = paste0(event_date, " (asp)"))

  pneumonia_last_year_to_date <- bind_rows(pneumonia_to_date, pneumonia_data_last_year, asp_pne_last_year_to_date) |>
  mutate(event_date_char = as.character(event_date))

} else if (nrow(aspiration_to_date) == 0 & nrow(aspiration_data_last_year) > 0) {

  asp_pne_last_year_to_date <- bind_rows(aspiration_data_last_year) |>
  mutate(event_date_char = paste0(event_date, " (asp)"))

  pneumonia_last_year_to_date <- bind_rows(pneumonia_to_date, pneumonia_data_last_year, asp_pne_last_year_to_date) |>
  mutate(event_date_char = as.character(event_date))

} else {

  pneumonia_last_year_to_date <- bind_rows(pneumonia_to_date, pneumonia_data_last_year) |>
  mutate(event_date_char = as.character(event_date))

}

#remove duplicate events
repeat_asp_pne2 <- anti_join(pneumonia_last_year_to_date |> select(-event_date_char),
                             asp_pne_last_year_to_date |> select(-event_date_char)) |> 
  mutate(
    event_date_char = as.character(event_date)
  )

repeat_asp_pne <- bind_rows(repeat_asp_pne2, asp_pne_last_year_to_date)

repeat_asp_pne_name <- as.data.frame(str_split(repeat_asp_pne$last_name, pattern = ",", simplify = TRUE))

repeat_asp_pne$last_name <- repeat_asp_pne_name$V1

#combine non-duplicated names with events
repeat_pneumonia <- inner_join(repeat_asp_pne, combined_names) |>
  mutate(name = paste(first_name, last_name, sep = " "))

repeat_pneumonia_person <- repeat_pneumonia |>
  count(name, name = "events_person")

repeat_pneumonia_date <- repeat_pneumonia |>
  group_by(name) |>
  arrange(event_date) |>
  mutate(conc_date = str_flatten(event_date_char, collapse = ", ")) |>
  select(first_name, last_name, dob, conc_date)

repeat_pneumonia_person_date <- left_join(repeat_pneumonia_person,
                                          repeat_pneumonia_date) |>
  filter(events_person > 1)

repeat_pneumonia_person_date |>
  distinct(name, events_person, conc_date) |>
  arrange(desc(events_person)) |>
  gt(rowname_col = "name") |>
  cols_label(events_person = "Events",
             conc_date = "Dates") 

```

